---
title: "Lab 03"
subtitle: 'Distances and the Border Zone'
author: "[Hope Wilcox](https://hopewilcox.github.io)"
output: 
  html_document:
    theme: flatly
---
# Libraries
```{r, message=F, waring = F}
# SPDS
library(tidyverse)
library(sf)
library(units)
library(dplyr)

# Data
library(USAboundaries)
library(rnaturalearth)

# Visualization
library(gghighlight)
library(ggrepel)
library(knitr)
library(ggthemes)
```
# Question 1:

## 1.1 Define a Projection
```{r, message=F, waring = F}
eqdc = '+proj=eqdc +lat_0=40 +lon_0=-96 +lat_1=20 +lat_2=60 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs'
```

### 

## 1.2 Get USA state boundaries


Access state boundaries

Filter data to only continental USA
```{r, message=F, waring = F}
conus = USAboundaries::us_states() %>%
  filter(!state_name %in% c("Puerto Rico",
                            "Alaska",
                            "Hawaii")) %>% 
  st_transform(5070)
```


## 1.3 Get country boundaries for Mexico, the USA, and Canada


World boundaries

Make sure data is in simple features format.

```{r, message=F, waring = F}
countries_sf = st_as_sf(countries110) %>% st_transform(5070)
```

Filter to Mexico, USA, Canada
```{r, message=F, waring = F}
mex_can_us = countries_sf %>% 
  filter(admin %in% c("Canada", "United States of America", "Mexico"))
  
```

## 1.4 Get city locations from the CSV file
```{r, message=F, waring = F}
cities = cities <- read_csv("../data/uscities.csv") %>% 
  st_as_sf(coords=c("lng","lat"), crs=4326) %>% 
   filter(!state_name %in% c("Puerto Rico",
                            "Alaska",
                            "Hawaii")) %>%  st_transform(5070)
```
# Question 2:
## 2.1 Distance to USA Border (coastline or national) (km)
```{r, message=F, waring = F}

conus_resolved = st_union(conus) %>%
  st_cast("MULTILINESTRING")


cities = cities %>% 
  mutate(distance_to_us_border = st_distance(cities, conus_resolved),
         distance_to_us_border = units::set_units(distance_to_us_border, "km"),
         distance_to_us_border = units::drop_units(distance_to_us_border))

farthest5_from_border = cities %>% 
  slice_max(distance_to_us_border, n=5) %>% 
  select(city, state_name, distance_to_us_border) %>% 
  st_drop_geometry()

knitr::kable(farthest5_from_border,
             caption = "5 US Cities Farthest from Border",
             col.names = c("City", "State", "Distance to Border"))

```

## 2.2 Distance to States (km)
```{r, message=F, waring = F}

conus_preserved = st_combine(conus) %>%
  st_cast("MULTILINESTRING")

cities = cities %>%
mutate(distance_to_states = st_distance(cities, conus_preserved),
     distance_to_states = units::set_units(distance_to_states, "km"),
     distance_to_states = units::drop_units(distance_to_states))



farthest_from_state = cities %>%
  slice_max(distance_to_states, n=5) %>%
  select(city, state_name, distance_to_states) %>%
  st_drop_geometry()

knitr::kable(farthest_from_state,
           caption = "5 US Cities Farthest from State Border",
            col.names = c("City", "State", "Distance to state border"))
```

## Distance to Mexico (km)
```{r, message=F, waring = F}
mexico = countries_sf %>% 
  filter(admin %in% c("Mexico")) 
  
cities = cities %>% 
  mutate(distance_to_mexico = st_distance(cities, mexico),
         distance_to_mexico = units::set_units(distance_to_mexico, "km"),
         distance_to_mexico = units::drop_units(distance_to_mexico))


farthest_from_mexico = cities %>% 
  slice_max(distance_to_mexico, n=5) %>% 
  select(city, state_name, distance_to_mexico) %>% 
  st_drop_geometry()

knitr::kable(farthest_from_mexico,
             caption = "5 US Cities Farthest from Mexico",
             col.names = c("City", "State", "Distance to Mexico"))
```


## Distance to Canada (km)
```{r, message=F, waring = F}
canada = countries_sf %>% 
  filter(admin %in% c("Canada"))


cities = cities %>% 
  mutate(distance_to_canada = st_distance(cities, canada),
         distance_to_canada = units::set_units(distance_to_canada, "km"),
         distance_to_canada = units::drop_units(distance_to_canada))

farthest_from_canada = cities %>% 
  slice_max(distance_to_canada, n=5) %>% 
  select(city, state_name, distance_to_canada) %>% 
  st_drop_geometry()

knitr::kable(farthest_from_canada,
             caption = "5 US Cities Farthest from Canada",
             col.names = c("City", "State", "Distance to Canada"))

```

# Question 3

## 3.1 Data
```{r, message=F, waring = F}
largest10_cities =cities %>% 
            slice_max(population, n=10)

cityfromborder = cities %>% 
  arrange(-distance_to_states) %>% 
  slice_max(distance_to_states, n = 5)

ggplot()+
  geom_sf(data = mex_can_us, color = "black") +
  geom_sf(data = conus, color = "black") +
  geom_sf(data = largest10_cities, color = "red", size = 2)+
  ggrepel::geom_label_repel(
    data = largest10_cities,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 2
  )

```

## 3.2 City Distance from the Border
```{r, message=F, waring = F, eval=F}
ggplot() + 
  geom_sf(data = conus) +
  geom_sf(data = cities, aes(col = distance_to_us_border), size = .1) +
  scale_color_gradient(low = "gray", high = "red") +
  geom_sf(data = farthest5_from_border, col = "blue", size = 2)+
    ggrepel::geom_label_repel(
      data = farthest5_from_border,
      aes(label = city, geometry = geometry),
          stat = "sf_coordinates",
          size = 2)
          
```

## 3.3 City Distance from Nearest State
```{r, message=F, waring = F, eval = F}
ggplot()+
geom_sf(data = farthest5_from_border, color = "green")+
geom_sf(data = farthest_from_state)+
geom_sf(data = farthest_from_state, aes(col = as.numeric(dist)), size=.1) +
ggrepel::geom_label_repel(data = farthest5_from_border,
                        aes(label = city, geometry=geometry),
                       stat = "sf_coordinates",
                      size = 2)
```



# Question 4 Real World Application
## 4.1 Quantifing Border Zone
```{r, message=F, waring = F, eval = F}
mex_cities_160km = cities %>% 
  filter(distance_to_mexico<160)

can_cities_160km = cities %>% 
  filter(distance_to_canada<160)

to_border

cities = cities %>% 
  mutate(mex_cities_160km = distance_to_mexico<160) %>% 
  mutate(can_cities_160km = distance_to_canada<160)


```


## 4.2 Mapping Border Zone
```{r, message=F, waring = F, eval = F}
ggplot()+
  geom_sf(data = mex_can_us, color = "black")+
  geom_sf(data = mex_cities_160km, color = "red")+
  geom_sf(data = can_cities_160km, color = "blue")
  
```








