---
title: "Geography 176A"
author: "[Hope Wilcox](https://hopewilcox.github.io)"
subtitle: 'Lab 03: Distances and the Border Zone'
output: 
  html_document:
    theme: flatly
---
# Libraries
```{r, results='hide'}
# SPDS
library(tidyverse)
library(sf)
library(units)
library(dplyr)

# Data
library(USAboundaries)
library(rnaturalearth)

# Visualization
library(gghighlight)
library(ggrepel)
library(knitr)
library(ggthemes)
```
# Question 1:

## 1.1 Define a Projection
```{r, eval=FALSE}
eqdc = '+proj=eqdc +lat_0=40 +lon_0=-96 +lat_1=20 +lat_2=60 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs'
```

### 

## 1.2 Get USA state boundaries


Access state boundaries
```{r, eval=FALSE}
USAboundaries::us_states(resolution = "low")
```
Filter data to only continental USA
```{r}
conus = USAboundaries::us_states() %>%
  filter(!state_name %in% c("Puerto Rico",
                            "Alaska",
                            "Hawaii"))
```


## 1.3 Get country boundaries for Mexico, the USA, and Canada
```{r, eval=FALSE}
remotes::install_github("ropenscilabs/rnaturalearthdata")
```

World boundaries
```{r, eval=FALSE}
rnaturalearth::countries110
```
Make sure data is in simple features format.
```{r}
countries_sf = st_as_sf(countries110, coords = c("lng","lat"), crs = 4326)
```
Filter to Mexico, USA, Canada
```{r}
mex_can_us = countries_sf %>% 
  filter(admin %in% c("Canada", "United States of America", "Mexico"))
  
```

## 1.4 Get city locations from the CSV file
```{r, results='hide'}
cities = cities <- read_csv("../data/uscities.csv") %>% 
  st_as_sf(coords=c("lng","lat"), crs=4326) %>% 
   filter(!state_name %in% c("Puerto Rico",
                            "Alaska",
                            "Hawaii"))
```
# Question 2:
## 2.1 Distance to USA Border (coastline or national) (km)
```{r}

conus_resolved = st_union(conus) %>%
  st_cast("MULTILINESTRING")


cities = cities %>% 
  mutate(distance_to_us_border = st_distance(cities, conus_resolved),
         distance_to_us_border = units::set_units(distance_to_us_border, "km"),
         distance_to_us_border = units::drop_units(distance_to_us_border))

farthest5_from_border = cities %>% 
  slice_max(distance_to_us_border, n=5) %>% 
  select(city, state_name, distance_to_us_border) %>% 
  st_drop_geometry()

knitr::kable(farthest5_from_border,
             caption = "5 US Cities Farthest from Border",
             col.names = c("City", "State", "Distance to Border"))

```

## 2.2 Distance to States (km)
```{r}

conus_preserved = st_combine(conus) %>%
  st_cast("MULTILINESTRING")

#cities = cities %>% 
  #mutate(distance_to_states = st_distance(cities, conus_preserved),
   #      distance_to_states = units::set_units(distance_to_states, "km"),
    #     distance_to_states = units::drop_units(distance_to_states))



#farthest_from_state = cities %>% 
 # slice_max(distance_to_states, n=5) %>% 
  #select(city, state_name, distance_to_states) %>% 
  #st_drop_geometry()

#knitr::kable(farthest_from_state,
 #            caption = "5 US Cities Farthest from State Border",
  #           col.names = c("City", "State", "Distance to state border"))
```

## Distance to Mexico (km)
```{r}
mexico = countries_sf %>% 
  filter(admin %in% c("Mexico")) 
  


cities = cities %>% 
  mutate(distance_to_mexico = st_distance(cities, mexico),
         distance_to_mexico = units::set_units(distance_to_mexico, "km"),
         distance_to_mexico = units::drop_units(distance_to_mexico))


farthest_from_mexico = cities %>% 
  slice_max(distance_to_mexico, n=5) %>% 
  select(city, state_name, distance_to_mexico) %>% 
  st_drop_geometry()

knitr::kable(farthest_from_mexico,
             caption = "5 US Cities Farthest from Mexico",
             col.names = c("City", "State", "Distance to Mexico"))
```


## Distance to Canada (km)
```{r}
canada = countries_sf %>% 
  filter(admin %in% c("Canada"))


cities = cities %>% 
  mutate(distance_to_canada = st_distance(cities, canada),
         distance_to_canada = units::set_units(distance_to_canada, "km"),
         distance_to_canada = units::drop_units(distance_to_canada))

farthest_from_canada = cities %>% 
  slice_max(distance_to_canada, n=5) %>% 
  select(city, state_name, distance_to_canada) %>% 
  st_drop_geometry()

knitr::kable(farthest_from_canada,
             caption = "5 US Cities Farthest from Canada",
             col.names = c("City", "State", "Distance to Canada"))

```

# Question 3

## 3.1 Data
```{r}
largest10_cities =cities %>% 
            slice_max(population, n=10)

ggplot()+
  geom_sf(data = mex_can_us, color = "black") +
  geom_sf(data = conus, color = "black") +
  geom_sf(data = largest10_cities, color = "red") 

```

## 3.2 City Distance from the Border
```{r}

```

## 3.3 City Distance from Nearest State
```{r}
#ggplot()+
 # geom_sf(data = farthest5_from_border, color = "green")+
  #geom_sf(data = distance_to_states)+
  #geom_sf(data = distance_to_states, aes(col = as.numeric(dist)), size=.1)+
  #ggrepel::geom_label_repel(data = farthest5_from_border,
   #                         aes(label = city, geometry=geometry),
    #                        stat = "sf_coordinates",
     #                       size = 2)
```



# Question 4 Real World Application
## 4.1 Quantifing Border Zone
```{r}
mex_cities_160km = cities %>% 
  filter(distance_to_mexico<160)

can_cities_160km = cities %>% 
  filter(distance_to_canada<160)

cities = cities %>% 
  mutate(mex_cities_160km = distance_to_mexico<160) %>% 
  mutate(can_cities_160km = distance_to_canada<160)


```


## 4.2 Mapping Border Zone
```{r}
ggplot()+
  geom_sf(data = mex_can_us, color = "black")+
  geom_sf(data = mex_cities_160km, color = "red")+
  geom_sf(data = can_cities_160km, color = "blue")
  
```








